### 2017-06-28 manuscript.rmd
included Physalia tentilla GV comment

The most complex nematocyst batteries of Cnidaria can arguably be found among the siphonophores, hanging in regularly spaced tentacle side branches called tentilla. Most hydrozoans, including the clade that contains siphonophores, bear simple tentacles (tentacles with no side branches). Is is still an open question whether the most recent common ancestor (MRCA) of Siphonophora had simple or branched tentacles. The only two siphonophore genera lacking tentilla are *Physalia* and *Apolemia*, at each side of the earliest split in the siphonophore tree, with an immediate sister group with all members bearing tentilla (*Rhizophysidae* and *{Bargmannia, Diphyes}*) (FigXX). Figure X shows the evolution of this character on the current phylogeny, indicating a most-likely tentilla bearing MRCA, with two independent losses. However, this reconstruction bears some uncertainty on the state of the MRCA, suggesting a ~30% probability of 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and *{Bargmannia, Diphyes}*.

A key issue here is how we code for absence of tentilla, especially for the case of *Physalia physalis*. The tentacles of this species, when uncoiled, show very prominent, evenly spaced, 3D bands which contain all active and functionally arranged nematocysts used by the organism for prey capture. We suspect these may be reduced tentilla. So far we do not have enough tissue-specific gene expression and histological data to accurately resolve this question. However, when we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction change radically (Figure X), indicating there was a single loss of tentilla in the branch leading to *Apolemiidae*. The main implication of this scenario is the unambiguous tentilla-bearing state of the MRCA.

>>>>

The most complex nematocyst batteries of Cnidaria can arguably be found among the siphonophores, hanging in regularly spaced tentacle side branches called tentilla. Most hydrozoans, including the clade that contains siphonophores, bear simple tentacles (tentacles with no side branches). Is is still an open question whether the most recent common ancestor (MRCA) of Siphonophora had simple or branched tentacles. The only two siphonophore genera regarded as lacking tentilla are *Physalia* and *Apolemia*, at each side of the earliest split in the siphonophore tree, with an immediate sister group with all members bearing tentilla (*Rhizophysidae* and *{Bargmannia, Diphyes}*) (FigXX). Figure X shows the evolution of this character on the current phylogeny, indicating a most-likely tentilla bearing MRCA, with two independent losses. However, this reconstruction bears some uncertainty on the state of the MRCA, suggesting a ~30% probability of 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and *{Bargmannia, Diphyes}*.

A key issue here is how we code for absence of tentilla, especially for the case of *Physalia physalis*. The tentacles of this species, when uncoiled, show very prominent, evenly spaced, 3D buttons which contain all active and functionally arranged nematocysts used by the organism for prey capture. Siphonophore tentilla are complete diverticular brachings of the tentacle ectoderm, mesoglea, and gastrovascular canal (lined by endoderm). Hessinger & Ford 1988 (in the Biology of Nematocysts) described these bands as enclosing individual fluid-filled chambers connected by narrow channels to the tentacular canal, lined by endoderm. This suggests they are not just ectodermal swellings, but probably are reduced tentilla. When we code *Physalia physalis* as tentilla bearing, the results for the character reconstraction change to a more parsimonious scenario (Figure X), indicating there was a single loss of tentilla in the branch leading to *Apolemiidae*. This result suggests an unambiguous tentilla-bearing state of the MRCA.
